Summary of Late Winter2013 Production: ip\_diffim

### 2013-03-15 W13report.tex
Edits on the math background.

@@

@@

We perturbed the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4) pixels, and to randomize the offsets multiplied
each offset by a number pulled from a normal (0,1) distribution.  The
output RMS reported by RegisterTask was noted to track these offsets.

@@

>>>>

\section{Introduction}

The focus of this production is to ascertain the ability of the LSST
software stack to provide clean subtractions between a science image
and a reference image (referred to as the template image below).  The
science image will typically be a single--epoch {\tt calexp}, while
the template image will be a high signal--to--noise representation of
the field, typically a {\tt coadd}.

The metric we use to define ``clean'' is the nubmer of false positives
that are detected and measured on the image difference.  We calculate
several additional quality metrics during the image subtraction
process, and correlate them with the number of false positives, as a
means to predict the number of false positives we may expect
downstream from the image subtraction itself.

\subsection{Algorithm}

We establish the math behind the process below.

We assume that science image $S(x,y)$ can be modeled as a convolution
of the template image $T(x,y)$ by a single PSF--matching kernel
$K(u,v,x,y)$.  We assume that the kernel may be decomposed using a set
of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients
in front of each basis are determiend through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v,x,y) = \sum_i a_i(x,y) K_i(u,v)$.  The image difference is
then calculated through $D(x,y) = S(x,y) - T(x,y) \otimes K(u,v,x,y)$.

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's Psf, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ Psf_S(x,y,u,v)$ where $\circ$
denotes correlation.  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the variance.

In this production, we investigate two orders of operations for the
optimal filtering by the Psf.  In the first, we Psf--match the
template image to a pre--filtered science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ Psf_S(x,y,u,v) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y) \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ Psf_S(x,y,u,v)$ -- we more frequencly avoid the case
of psf--matching {\it deconvolution}, where the FWHM of $T$ is
narrower than $S$.  This pre--filtering increases the FWHM of the
image by $\sqrt{2}$.  The Psf matching kernels will need to be
correspondingly larger to account for the larger Psfs.  We are able to
run detection directly on this image.

In the second, which is the classical implementation of the problem,
we create a difference image which is then post--filtered with the Psf
for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ Psf_S(x,y,u,v)  \nonumber 
\end{eqnarray}

\subsection{Simulations}

\section{Implementation}

We use the ImageDiffereceTask to

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/wcs_rms.eps} 
\includegraphics[width=0.5\textwidth]{figures/wcs_offset.eps} \\
\caption{WCS}
\label{wcs_offsets}
\end{figure}

@@

We perturbed the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4) pixels, and to randomize the effects we
multiplied each offset by a number pulled from a normal (0,1)
distribution.  The output RMS reported by RegisterTask was noted to
track these offsets.

@@

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single KernelCandidate local difference
  image (visit=6866603, doPreFilter=True).  Columns indicate the
  Gaussian sigma of the basis, which is designed to have one Gaussian
  with the prescribed width, modified to fourth order.  Rows indicate
  the amplitude of the astrometric offset that was introducted into
  the Source positions before image-to-image registration.  \\
  In the {\it top} row, we see that even with a small astrometric
  offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image. \\
  The ability of the basis to compensate for bulk astrometric offsets
  is a function of the kernel width, and the kernel width is itself a
  function of the two input Psf FWHMs.  Thus there is a compliated
  dependence of the ability of the kernel to shift the centroids of
  stars.  Roughly, the basis may compensate for local astrometric residuals
  at a scale up to $\sqrt{\sigma_I^2 - \sigma_R^2}$.}
\label{kernel_offsets}
\end{figure}
